---
title: "402 HW"
author: "Ruiqi Zhang"
date: "2020/11/29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tm)
library(SnowballC)
library(lda)
library(topicmodels)
library(LDAvis)
library(dplyr)
library(stringi)
library(plyr)
library(foreign)
#library(xts)
library(tis)
library(jsonlite)
library(FNN)
library(hexbin)
library(RColorBrewer)
library(MASS)
library(ldatuning)
library(gofastr)
library(quantmod)
library(tseries)
library(foreign)
library(forecast)
library(MASS)
library(TTR)
library(vars)
library(readtext) 
library(tidyr) 
library(scales)
library(tinytex)
library(fitdistrplus)
library(rgl)
library(car)
library(zoo)
```

# 1

Acquire data first.

```{r cars}
#get data
data <- ts(read.csv("C://Users//zhangruiqi//Desktop//402 Macrotheory//HW 402//Dataset for Homework.csv",header=FALSE)[,1],start=1)
head(data)
```

(A)
Plot data

```{r pressure, echo=FALSE}
plot(data)
```

Test if stationary
```{r}
#ADF test
adf.test(data,k=1)
#kpss test
kpss.test(data)

```
The data is stationary. 

(B)
```{r}
#Using first 150 steps construct AR(1) model
data_150=data[1:150]

fit=arima(as.numeric(data_150) ,order = c(1,0,0))
fit
```

(C)Recursive prediction
```{r}
#Construct new dataframe
datapred<-data.frame(realdata=data,fitdata=numeric(200),error=numeric(200))

#Recursive prediction
for(i in 1:50){
# Each iteration construct a new AR(1) model using known and predicted data acquired
fit_recur=arima(as.numeric(datapred$realdata[1:150+i-1]),order = c(1,0,0))
#Point prediction for each iteration
datapred$fitdata[150+i]=as.numeric(fit_recur$coef[1])*as.numeric(datapred$realdata[150+i-1])+as.numeric(fit_recur$coef[2])
#Calculate error each iteration
datapred$error[150+i]=as.numeric(datapred$realdata[150+i])-as.numeric(datapred$fitdata[150+i])
}

#Calculate RMSE
errorpred=datapred$error[151:200]
RMSE=sqrt(mean((errorpred)^2))
RMSE
```
(D)
```{r}
#AR(2) model
fit2=arima(as.numeric(data_150) ,order = c(2,0,0))
fit2
```

```{r}
#Construct new dataframe2
datapred2<-data.frame(realdata=data,fitdata2=numeric(200),error2=numeric(200))

#Recursive prediction AR(2)
for(i in 1:50){
# Each iteration construct a AR(2) model using known and predicted data acquired
fit_recur2=arima(as.numeric(datapred2$realdata[1:150+i-1]),order = c(2,0,0))
#Point prediction for each iteration
datapred2$fitdata2[150+i]=as.numeric(fit_recur2$coef[1])*as.numeric(datapred2$realdata[150+i-1])+as.numeric(fit_recur2$coef[2])*as.numeric(datapred2$realdata[150+i-2])+as.numeric(fit_recur2$coef[3])
#Calculate error each iteration
datapred2$error2[150+i]=as.numeric(datapred2$realdata[150+i])-as.numeric(datapred2$fitdata2[150+i])
}

#Calculate RMSE
errorpred2=datapred2$error2[151:200]
RMSE2=sqrt(mean((errorpred2)^2))
RMSE2
```
RMSE for AR(1) model is 2.191, RMSE for AR(2) model is 2.194. AR(1) model is better. 

#2
```{r}
#construct GDP time series
GDPdata <- read.csv("C://Users//zhangruiqi//Desktop//402 Macrotheory//HW 402//GDPC1.csv",header = TRUE)
GDPdata =ts(GDPdata$GDPC1,start = 1950,frequency = 4)
head(GDPdata)
```

Take logs and plot the data. 
```{r}
logGDP=log(GDPdata)
plot(logGDP,main="Log US GDP data")

```
Take first difference of log-GDP data. 
```{r}
#log-first difference and log
lGDP_diff=diff(logGDP)
plot(lGDP_diff)
#lGDP_diff
```
Test if the log first difference data was stationary. 
```{r}
#adf test
adf.test(lGDP_diff)
#kpss test
kpss.test(lGDP_diff)
```

The data is stationary. 
There is autocorrelation in this dataset. Plot ACF and PACF for the data. 
```{r}
#plot acf
acf(lGDP_diff)
#plot pacf
pacf(lGDP_diff)
```

From the PACF plot, there is only one spike standing out. This seems like an AR(1) process. Use an AR(1) model. 
```{r}
#Construct an AR(1) model
gdpfit=arima(as.numeric(window(lGDP_diff, start=1950, end=c(2011,4), frequency=4)) ,order = c(1,0,0))
gdpfit

```
Recursive prediction. 
```{r}
#Construct new data frame for training set
datapredgdp<-data.frame(realdata=lGDP_diff,fitdata=numeric(length(lGDP_diff)),error=numeric(length(lGDP_diff)))
datapredgdp$fitdata[1:248]=datapredgdp$realdata[1:248]

#Recursive prediction
for(i in 1:32){
# Each iteration update a new AR(1) model using the latest actual data acquired
recur_gdp=arima(as.numeric(datapredgdp$realdata[1:248+i-1]),order = c(1,0,0))
#Point prediction for each iteration
datapredgdp$fitdata[248+i]=as.numeric(recur_gdp$coef[1])*as.numeric(datapredgdp$realdata[248+i-1])+as.numeric(recur_gdp$coef[2])
#Calculate error each iteration
datapredgdp$error[248+i]=as.numeric(datapredgdp$realdata[248+i])-as.numeric(datapredgdp$fitdata[248+i])
}


#datapredgdp
datapredgdp$realdata[1]
```
```{r}
#time dummy
t=seq(1950.25,2020,length=length(GDPdata))
#Construct predicted GDP
logGDP_pre=numeric(length(logGDP))
logGDP_pre[1]=logGDP[1]
for (i in 1:280) {
  logGDP_pre[1+i]=logGDP_pre[i]+datapredgdp$fitdata[i]
}
GDP_pre=exp(logGDP_pre)
GDP_pre=ts(GDP_pre,start=c(1950,1),frequency = 4)
```

```{r}
#Plot actual data and predicted data
plot(GDPdata,xlim=c(2012,2020),ylim=c(15000,22000))
lines(t,GDP_pre,col="red")
```

```{r}
#Calculate pretiction error
errorpredgdp=GDP_pre[249:281]-GDPdata[249:281]
#calculate RMSE
RMSEgdp=sqrt(mean((errorpredgdp)^2))
RMSEgdp
```

